### 2018-05-11 algorithm.tex
analytical solution under free trade

\documentclass{article}
\usepackage{amsmath}
\begin{document}
Solving a simple Armington-endowment model.

Let $Q_n = A_NL_n$, with $L_n$ completely inelastic. Demand is CES:
\[
C_n^{1-1/\theta} = \sum_m Q_m^{1-1/\theta}.
\]
Trade costs are $\tau_{ij}$ and we denote $\kappa_{ij} = \tau_{ij}^{-\theta}$.

This model has the following equilibrium equations.
\[
d_{nm} = 
	\frac
	{(\tau_{nm}w_m)^{1-\theta}}
	{\sum_k (\tau_{nk} w_k)^{1-\theta}}
\]
\[
w_m \cdot L_m = \sum_n d_{nm} w_n \cdot L_n
\]
where $w_m$ is the unit cost of labor $L_m$.
\[
w_m L_m = w_m^{1-\theta}
\sum_n 	\frac
	{\tau_{nm}^{1-\theta}}
	{\sum_k (\tau_{nk} w_k)^{1-\theta}}
 w_n L_n
\]
Let $\Phi_n = \sum_k (\tau_{nk} w_k)^{1-\theta}$ 
\[
w_m^\theta L_m = 
\sum_n 	
	{\tau_{nm}^{1-\theta}}
 w_n L_n/\Phi_n.
\]
This leads to a better iteration: Pick $w_n$ and calculate corresponding $\Phi_n$. Calculate new $w_n$ from LHS.

\begin{enumerate}
	\item Pick $\alpha_m^{(0)} = w_mL_m/\Phi_m$ from some previous equilibrium (or data).
	\item Solve for
	\[
	w_m^{(k+1)} = L_m^{-1/\theta} 
		\left[
		\sum_n \alpha_n^{(k)} \tau_{nm}^{1-\theta}
		\right]^{1/\theta}
	\]
	\item Set 
	\[
	\alpha_m^{(k+1)} = w_m^{(k)} L_m/\Phi_m(w^{(k)})
	\]
	and repeat.
\end{enumerate}
This should converge fast, in 3-4 steps, because (i) we start from a close guess, (ii) $\theta$ is large so that $1/\theta$ is small, resulting in dampening, (iii) $\alpha$ has limited impact on wages, only due to destination-country variation in $\tau$, dampening feedback.

\section{Full model}
Introduce a new variable for the factory-gate price of intermediate goods,
\begin{equation}\label{eq:varrho}
	\varrho_{njt} = \xi B_j T_{nj}^{-1/\theta} A_{njt}^{-1} w_{njt}^{\beta_j} \prod_{k=1}^J P_{nkt}^{\gamma_{kj}}.
\end{equation}
With this variable, we can write the price of the final good as
\begin{equation}\label{eq:price}
	P_{mjt} = \left[
		\sum_{n=1}^N (\varrho_{njt}/\kappa_{mnjt})^{-\theta}
		\right]^{-1/\theta}.
\end{equation}
This is a standard CES price index. We can write import shares as
\begin{equation}\label{eq:import_share}
	d_{mnjt} = \left[
	\frac
	{\varrho_{njt}/\kappa_{mnjt}}
	{P_{mjt}}
	\right]^{-\theta}.
\end{equation}
Wages can be expressed by inverting \eqref{eq:varrho},
\begin{equation}\label{eq:wage}
	w_{njt} = (\xi B_j)^{-1/\beta_j}
	\varrho_{njt}^{1/\beta_j}  
	T_{nj}^{1/(\beta_j\theta)}A_{njt}^{1/\beta_j} 
	\prod_{k=1}^J P_{nkt}^{-\gamma_{kj}/\beta_j}.
\end{equation}

Taking the market clearing condition,
\begin{equation}
	R_{njt} = \sum_{m=1}^N
		d_{mnjt}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\end{equation}
\[
	R_{njt}\varrho_{njt}^{\theta} = \sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right],
\]
\begin{equation}\label{eq:market_clearing}
	\varrho_{njt} = R_{njt}^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		(\kappa_{mnjt} P_{mjt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right]
	\right\}^{1/\theta}.
\end{equation}
Here both $R_{njt}$ and $\varrho_{njt}$ depend on equilibrium prices. But starting from a guess for $R$ (say, the autarky equilibrium), we can easily compute a shadow price $\varrho$ supporting that allocation. 
\begin{enumerate}
	\item Set $R_{njt}^{(0)}$ and $P_{njt}^{(0)}$ to the autarky equilibrium value or other suitable starting value.
	\item Given $R_{njt}^{(k)}$ and $P_{njt}^{(k)}$, solve for $\varrho_{njt}^{(k+1)}$ using equation \eqref{eq:market_clearing}.
	\item Solve for $w_{njt}^{(k+1)}$ and $P_{njt}^{(k+1)}$ using \eqref{eq:wage} and \eqref{eq:price}.
	\item Calculate $R_{njt}^{(k+1)} = w_{njt}^{(k+1)}L_{njt}/\beta_j$.
	\item Go back to Step 2 and repeat until convergence.
\end{enumerate}
Typically, this converges in 4--6 steps. This might even be a contraction mapping.

\begin{equation*}
	\varrho_{njt}/P_{njt} = 
	\left\{
	\sum_{m=1}^N
		(\kappa_{nmjt} P_{mjt}/P_{njt})^{\theta}
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt}/R_{njt} - S_{mt}/R_{njt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}/R_{njt}
		\right]
	\right\}^{1/\theta}.
\end{equation*}

\subsection{Suitable starting values}
Suppose that trade is free, $\kappa_{nmjt}\equiv 1$. Then $P_{njt}\equiv P_{jt}$ and
\begin{equation*}
	\varrho_{njt}^{(0)} = 
	P_{jt}
	R_{njt}^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		\left[
			\alpha_{jt}
			\left(
				\sum_{k=1}^J\beta_k R_{mkt} - S_{mt}
			\right)
			+ \sum_{k=1}^J\gamma_{jk}R_{mkt}
		\right]
	\right\}^{1/\theta}.
\end{equation*}
Let $R_{wjt}$ denote the world revenue in sector $j$.
\begin{equation*}
	\varrho_{njt}^{(0)} = 
	P_{jt}
	R_{njt}^{-1/\theta}
		\left[
				\sum_{k=1}^J
				(\alpha_{jt}\beta_k + \gamma_{jk}) R_{wkt}
		\right]^{1/\theta}.
\end{equation*}
This only determines $\varrho_{njt}/P_{jt}$, what determines $P_{jt}$ relative to other sectors?\begin{equation*}
	\varrho_{njt}/P_{jt} = 
	\beta_j^{1/\theta}
	(w_{njt}L_{njt})^{-1/\theta}
		\left[
				\sum_{k=1}^J
				(\alpha_{jt} + \gamma_{jk}/\beta_k) 
				\sum_{m=1}^N w_{mkt}L_{mkt}
		\right]^{1/\theta}.
\end{equation*}
Sum across $n$s,
\begin{equation*}
	(\varrho_{njt}/P_{jt})^{-\theta} = 
	\beta_j^{-1}
	(w_{njt}L_{njt})
		\left[
				\sum_{k=1}^J
				(\alpha_{jt} + \gamma_{jk}/\beta_k) 
				\sum_{m=1}^N w_{mkt}L_{mkt}
		\right]^{-1}.
\end{equation*}
\begin{equation*}
	w_{wjt}L_{wjt}
	= \beta_j
				\sum_{k=1}^J
				(\alpha_{jt} + \gamma_{jk}/\beta_k) 
				w_{wkt}L_{wkt}.
\end{equation*}
\begin{equation*}
	R_{wjt}
	=
			\sum_{k=1}^J
			(\alpha_{jt}\beta_k + \gamma_{jk}) 
			R_{wkt}.
\end{equation*}
We can solve this for eigenvalue equation for $R_{wjt}$ without knowing anything about equilibrium prices, conditional a global numeraire.

Write the log of equation \eqref{eq:varrho} in matrix notation, and recognizing that under free trade, $P$ does not depend on $n$,
\[
\ln\mathbf \varrho_{nt} = \ln\xi
	+ \ln\mathbf B
	- \ln\mathbf A_{nt}
	+ \text{diag}(\beta) \ln\mathbf w_{nt}
	+ \Gamma \ln\mathbf P_{t}  
\]
\[
\ln\mathbf \varrho_{nt} - \ln\mathbf P_{t} 
 = \frac1\theta\ln\mathbf\beta
 	-\frac1\theta(\ln \mathbf w_{nt}+ \ln \mathbf L_{nt})
 	+\frac1\theta \ln \mathbf E_{wt}
\]
\[
\frac1\theta\ln\mathbf\beta
 	-\frac1\theta(\ln \mathbf w_{nt}+ \ln \mathbf L_{nt})
 	+\frac1\theta \ln \mathbf E_{wt}
=
\ln\xi
	+ \ln\mathbf B
	- \ln\mathbf A_{nt}
	+ \text{diag}(\beta) \ln\mathbf w_{nt}
	+ (\Gamma-\mathbf I) \ln\mathbf P_{t}
\]
\[
\ln\mathbf\beta
 	-\ln \mathbf L_{nt}
 	+ \ln \mathbf E_{wt}
	+ \theta\ln\mathbf A_{nt}
	-\theta\ln\xi
	- \theta\ln\mathbf B
	+ \theta(\mathbf I-\Gamma) \ln\mathbf P_{t}
=
	 \text{diag}(1+\beta\theta) \ln\mathbf w_{nt}
\]
Guess a $P$. Solve above for $w$. Then solve for next $P$.
\subsubsection{General problem for free trade}
Pricing equation:
\[
\ln\mathbf \varrho_{nt}-\ln\mathbf P_{t} = 
\ln\xi
	+ \ln\mathbf B
	- \ln\mathbf A_{nt}
+\text{diag}(\beta)(\ln\mathbf w_{nt}-\ln\mathbf P_{t})
+ \mathbf C \ln\mathbf P_{t}  
\]
The last matrix $\mathbf C = \Gamma+\text{diag}(\beta)-\mathbf I$ has the property $\mathbf C\mathbf 1=\mathbf 0$ so this equation is hom(0) in prices.

Import share equation:
\[
\ln\mathbf \varrho_{nt} - \ln\mathbf P_{t} 
 = 
 	+\frac1\theta(\ln\mathbf\beta+\ln \mathbf E_{wt}-\ln \mathbf w_{nt}- \ln \mathbf L_{nt})
\]
This is also hom(0) in prices. World expenditure $E_{wt}$ can be precomputed up to a choice of numeraire.

CES price index:
\[
\ln\mathbf P_{t} = -\frac1\theta
 \ln \left[
 	\sum_{n=1}^N \exp(-\theta \ln \varrho_{nt})
 \right]
\]
\[
\mathbf 1 =  
 	\sum_{n=1}^N \exp[-\theta (\ln \varrho_{nt}-\ln\mathbf P_{t})]
\]
Again, hom(0) in prices. Substituting in the import share equation,
\[
\mathbf 1 =  
 	\sum_{n=1}^N \exp(\ln\mathbf w_{nt}-\ln\mathbf E_{wt}-\theta\mathbf b_2).
\]
This just says that the import shares add up to 1. This suggests that we can limit the search to import shares on the simplex. Let $\mathbf d_{nt}$ denote the import share vector.
\[
\ln\mathbf \varrho_{nt} - \ln\mathbf P_{t} 
 = -\frac1\theta \ln\mathbf d_{nt}
\]
Wages are
\[
\ln\mathbf w_{nt} = \ln\beta+\ln\mathbf d_{nt}+\ln\mathbf E_{wt}-\ln\mathbf L_{nt}
\]
so that
\[
-\frac1\theta \ln\mathbf d_{nt} = 
\ln\xi
	+ \ln\mathbf B
	- \ln\mathbf A_{nt}
+
\text{diag}(\beta)(\ln\beta+\ln\mathbf d_{nt}-\ln\mathbf L_{nt}+\ln\mathbf E_{wt})
- (\mathbf I-\Gamma) \ln\mathbf P_{t}  
\]
\[
\text{diag}(\beta+1/\theta) \ln\mathbf d_{nt} = 
-\ln\xi
	- \ln\mathbf B
	+ \ln\mathbf A_{nt}
+
\text{diag}(\beta)(\ln\mathbf L_{nt}-\ln\mathbf E_{wt})
+ (\mathbf I-\Gamma) \ln\mathbf P_{t}  
\]
This holds for all $n$ and $t$. In particular, it also holds for country $0$, which we can subtract:
\[
\text{diag}(\beta+1/\theta) (\ln\mathbf d_{nt}-\ln\mathbf d_{0t}) = 
	\ln\mathbf A_{nt}-\ln\mathbf A_{0t}
+
\text{diag}(\beta)(\ln\mathbf L_{nt}-\ln\mathbf L_{0t})
\]
\[
(\ln\mathbf d_{nt}-\ln\mathbf d_{0t}) = 
	\text{diag}(\theta/(1+\beta\theta)) (\ln\mathbf A_{nt}-\ln\mathbf A_{0t})
+
\text{diag}(\beta\theta/(1+\beta\theta))(\ln\mathbf L_{nt}-\ln\mathbf L_{0t})
\]
This suggests the following solution for import shares,
\[
d_{njt} = c_{jt}A_{njt}^{\theta/(1+\beta_j\theta)}L_{njt}^{\beta_j\theta/(1+\beta_j\theta)},
\]
where the $c_{jt}$ constant is such that import shares add up to 1 for all $j$ and all $t$.

Given $d_{njt}$ and the precomputed $E_{wjt}$, we can calculate wages in all sectors. Prices can be calculated from
\[
\ln\mathbf P_{t} = 
(\mathbf I-\Gamma)^{-1}[\text{diag}(\beta+1/\theta) \ln\mathbf d_{0t}  
+\ln\xi
	+ \ln\mathbf B
	- \ln\mathbf A_{0t}
+
\text{diag}(\beta)(\ln\mathbf E_{wt}-\ln\mathbf L_{0t})]
 .  
\]

\subsubsection{Autarky}
Suppose that each country is in autarky, $\kappa_{nmjt}\equiv 0$ for all $n\neq m$ and 1 otherwise, and $S_{nt}\equiv 0$. Then $P_{njt}=\varrho_{njt}$ and 
\begin{equation*}
	R_{njt} = 			
		\sum_{k=1}^J(\alpha_{jt}\beta_k+\gamma_{jk}) R_{nkt}. 
\end{equation*}
Solving the eigenvalue equation
\[
\mathbf R_{nt} = (\mathrm\alpha_t \mathbf \beta'+\Gamma)\mathbf R_{nt}
\]
pins down $R_{njt}$ up to a scale parameter $\omega_{nt}$ and we can easily solve for wages, again up to the same scale parameter. Note that this only depends on fixed parameters, so we can pre-calculate autarky wages.

$\omega_{nt}R_{njt}$, $\omega_{nt}w_{njt}$

Write the log of equation \eqref{eq:varrho} in matrix notation and recognizing $\varrho_{njt}\equiv P_{njt}$ in autarky,
\[
\ln\mathbf P_{nt} = \ln\xi
	+ \ln\mathbf B
	- \ln\mathbf A_{nt}
	+ \text{diag}(\beta) \ln\mathbf w_{nt}
	+ \Gamma \ln\mathbf P_{nt}  
\]
\[
\ln\mathbf P_{nt} = 
(\mathbf I-\Gamma)^{-1}
	[
	\ln\xi
	+ \ln\mathbf B
	- \ln\mathbf A_{nt}
	+ \text{diag}(\beta) \ln\mathbf w_{nt}
	]
\]
This gives us prices up to the same scale parameter. Again, everything can be precalculated and the only part that varies across simulations is
\[
\Delta \ln\mathbf P_{nt}
-(\mathbf I-\Gamma)^{-1}
	\Delta\ln\mathbf A_{nt}.
\]
We still need to find suitable starting values for $\omega_{nt}$. We can use the overall GDP shares in the data 
\[
\omega_{nt} = c\cdot Y_{nt}^{-1/\theta}.
\]
This can also be precalculated.

\end{document}